---
title: "Assignment 7: GLMs (Linear Regressios, ANOVA, & t-tests)"
author: "Miaojun Pang"
date: "Spring 2024"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on generalized linear models. 

## Directions
1. Rename this file `<FirstLast>_A07_GLMs.Rmd` (replacing `<FirstLast>` with your first and last name).
2. Change "Student Name" on line 3 (above) with your name.
3. Work through the steps, **creating code and output** that fulfill each instruction.
4. Be sure to **answer the questions** in this assignment document.
5. When you have completed the assignment, **Knit** the text and code into a single PDF file.


## Set up your session 
1. Set up your session. Check your working directory. Load the tidyverse, agricolae and other needed packages. Import the *raw* NTL-LTER raw data file for chemistry/physics (`NTL-LTER_Lake_ChemistryPhysics_Raw.csv`). Set date columns to date objects.

2. Build a ggplot theme and set it as your default theme.

```{r setup2}
#1

getwd()
library(tidyverse)

NTL_LTER<- read.csv("/Users/mpang/OneDrive - Duke University/Desktop/Assignments/NTL-LTER_Lake_ChemistryPhysics_Raw.csv")
NTL_LTER$sampledate<- as.Date(NTL_LTER$sampledate,format = "%m/%d/%y")

#2

EDA_Theme<- theme_classic(base_size = 10) + theme(axis.title = element_text(color = "red"),
                                                  legend.position = "bottom")

print(EDA_Theme)

```

## Simple regression
Our first research question is: Does mean lake temperature recorded during July change with depth across all lakes?

3. State the null and alternative hypotheses for this question:
> Answer:
H0: There are no difference in mean lake temperature recorded during July change with depth across all lakes.
Ha: There is a difference in mean lake temperature recorded during July change with depth across all lakes.


4.  Wrangle your NTL-LTER dataset with a pipe function so that the records meet the following criteria: 
 * Only dates in July. 
 * Only the columns: `lakename`, `year4`, `daynum`, `depth`, `temperature_C`
 * Only complete cases (i.e., remove NAs)

5. Visualize the relationship among the two continuous variables with a scatter plot of temperature by depth. Add a smoothed line showing the linear model, and limit temperature values from 0 to 35 °C. Make this plot look pretty and easy to read.

```{r scatterplot}
#4
NTL_LTER_New <- NTL_LTER%>%
  separate(sampledate, c("Year", "Month",  "Day"), sep = "-")%>%
  filter(Month == "07")%>%
  select(`lakename`, `year4`, `daynum`, `depth`, `temperature_C`)
  
#5
library(ggplot2)
Temperaturebydepth<-   NTL_LTER_New%>%
  ggplot(aes(x=depth, y=temperature_C))+
  geom_point(aes(color=lakename))+
  geom_smooth(method = lm, col="red")+
  labs(x="Depth", y="Temperature (C)")+
  ylim(0,35)+
  theme_bw()+
  theme(legend.position = "bottom")
print(Temperaturebydepth)

```


6. Interpret the figure. What does it suggest with regards to the response of temperature to depth? Do the distribution of points suggest about anything about the linearity of this trend?

> Answer: The figure shows that the temperature decreases as the depth increases. And the distribution of points sugguest the temperature is close related to the linearity of this trend.


7. Perform a linear regression to test the relationship and display the results.

```{r linear.regression}
#7
temperature_linear<- lm(data= NTL_LTER_New,temperature_C~ depth)

summary(temperature_linear)

```


8. Interpret your model results in words. Include how much of the variability in temperature is explained by changes in depth, the degrees of freedom on which this finding is based, and the statistical significance of the result. Also mention how much temperature is predicted to change for every 1m change in depth. 

> Answer:
On the summary, the model results have strong relationsip with the depth, there are over 73% of the temperature variable is explained by the changes in lake depth. The residual standard error is 3.835 on 9726 degree of freedom. As the p value is very small, which means there is significant in this case. 

---

## Multiple regression
Let's tackle a similar question from a different approach. Here, we want to explore what might the best set of predictors for lake temperature in July across the monitoring period at the North Temperate Lakes LTER. 


9. Run an AIC to determine what set of explanatory variables (year4, daynum, depth) is best suited to predict temperature.

10. Run a multiple regression on the recommended set of variables. 

```{r temperature.model}
#9

NTLregression<- lm(data = NTL_LTER_New,
                   temperature_C~depth+year4+daynum)
step(NTLregression)

#10
NTLmultipleRegression<-
  lm(data = NTL_LTER_New, 
     temperature_C~depth+year4+daynum)

summary(NTLmultipleRegression)

```

11. What is the final set of explanatory variables that the AIC method suggests we use to predict temperature in our multiple regression? How much of the observed variance does this model explain? Is this an improvement over the model using only depth as the explanatory variable?

> Answer: 
The AIC method suggests that we use depth, year, and daynum to predict temperature. The multiple regression model shows that the R^2 value improved to a value of 0.7411. This means that the variables listed accounts for 74.11% of the change in temperature.



---
## Analysis of Variance

12. Now we want to see whether the different lakes have, on average, different temperatures in the month of July. Run an ANOVA test to complete this analysis. (No need to test assumptions of normality or similar variances.) Create two sets of models: one expressed as an ANOVA models and another expressed as a linear model (as done in our lessons).

```{r anova.model}
#12
laketempe_New<-NTL_LTER_New%>%
  group_by(lakename)%>%
  summarise(meantemp = mean(temperature_C))

laketempANOVA<- aov(data=NTL_LTER_New, temperature_C~lakename)
summary(laketempANOVA)

laketempANOVA2<- lm(data = NTL_LTER_New, temperature_C~lakename)
summary(laketempANOVA2)

```

13. Is there a significant difference in mean temperature among the lakes? Report your findings. 

> Answer: 
Yes, there is a significant differenc ein mean temperature among the lakes, as the p value shows, is smaller than 2.2e-16, with multiple r squred: 0.03953 and adjusted r squared: 0.03874.



14. Create a graph that depicts temperature by depth, with a separate color for each lake. Add a geom_smooth (method = "lm", se = FALSE) for each lake. Make your points 50 % transparent. Adjust your y axis limits to go from 0 to 35 degrees. Clean up your graph to make it pretty. 

```{r scatterplot.2}
#14.

tempbydepthLakes<-NTL_LTER_New%>%
  ggplot(aes(x=depth, y=temperature_C))+
  geom_point(aes(color=lakename), alpha=.5, size=0.5)+
  geom_smooth(aes(group=lakename, color=lakename),
    method = lm, size=0.4, se= FALSE)+
  ylim(0,35)+
  labs(x="Depth", y="Temperature (C)")+
  theme_bw()+
  theme(legend.position = "bottom")

print(tempbydepthLakes)

```

15. Use the Tukey's HSD test to determine which lakes have different means.

```{r tukey.test}
#15

TukeyHSD(laketempANOVA)

```

16.From the findings above, which lakes have the same mean temperature, statistically speaking, as Peter Lake? Does any lake have a mean temperature that is statistically distinct from all the other lakes?

>Answer:
From the Tukey multiple comparisons of means, there is 95% family-wise confidence level, which means they have significant relationships.
 

17. If we were just looking at Peter Lake and Paul Lake. What's another test we might explore to see whether they have distinct mean temperatures? 

>Answer: When we just looking at Peter Lake and Paul Lake, they are very similarand close. We could use two sample t test to explore to see whether they have distinct mean temperatures. 



18. Wrangle the July data to include only records for Crampton Lake and Ward Lake. Run the two-sample T-test on these data to determine whether their July temperature are same or different. What does the test say? Are the mean temperatures for the lakes equal? Does that match you answer for part 16?

```{r t.test}
NTL_LTER_New_TEST<- NTL_LTER_New %>%
  filter(lakename %in% c("Crampton Lake", "Ward Lake"))

NTL_TWOSAMPLE<- t.test(temperature_C ~ lakename, NTL_LTER_New_TEST)
print(NTL_TWOSAMPLE)
```

>Answer: The two-sample t test shows the p-value is 0.26. The mean in group 
Crampton Lake is 15.35, while the mean in group Ward Lake is 14.46. The mean 
temperatures for the lakes is not equal to 0. 
