---
title: "Assignment 10: Data Scraping"
author: "Miaojun Pang"
output: pdf_document
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on data scraping. 

## Directions
1. Rename this file `<FirstLast>_A10_DataScraping.Rmd` (replacing `<FirstLast>` with your first and last name).
2. Change "Student Name" on line 3 (above) with your name.
3. Work through the steps, **creating code and output** that fulfill each instruction.
4. Be sure your code is tidy; use line breaks to ensure your code fits in the knitted output.
5. Be sure to **answer the questions** in this assignment document.
6. When you have completed the assignment, **Knit** the text and code into a single PDF file.


## Set up 
1. Set up your session:

* Load the packages `tidyverse`, `rvest`, and any others you end up using.
* Check your working directory

```{r 1. Project Setup, message = FALSE}
#1 

getwd()
library(tidyverse)
library(rvest)
library(lubridate)

EDA_theme<- theme_classic(base_size = 10)+
            theme(axis.text = element_text (color = "black"),
                  legend.position = "bottom")
theme_set(EDA_theme)

```

2. We will be scraping data from the NC DEQs Local Water Supply Planning website, specifically the Durham's 2022 Municipal Local Water Supply Plan (LWSP): 
 * Navigate to https://www.ncwater.org/WUDC/app/LWSP/search.php
 * Scroll down and select the LWSP link next to Durham Municipality. 
 * Note the web address: <https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=03-32-010&year=2022>
 
Indicate this website as the as the URL to be scraped. (In other words, read the contents into an `rvest` webpage object.)

```{r set.the.scraping.website}
#2 

URL_data<- read_html('https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=03-32-010&year=2022')
URL_data

```

3. The data we want to collect are listed below:

* From the "1. System Information" section:
 * Water system name
 * PWSID
 * Ownership
 
* From the "3. Water Supply Sources" section:
 * Maximum Day Use (MGD) - for each month

In the code chunk below scrape these values, assigning them to four separate variables.

>HINT: The first value should be "Durham", the second "03-32-010", the third "Municipality", and the last should be a vector of 12 numeric values (represented as strings)".

```{r scrape.the.data}
#3 
water.system.name <- URL_data%>%
  html_nodes("div+ table tr:nth-child(1) td:nth-child(2)")%>%
  html_text()
water.system.name

pswid <- URL_data%>%
  html_nodes("td tr:nth-child(1) td:nth-child(5)")%>%
  html_text()
pswid

ownership <- URL_data%>%
  html_nodes("div+ table tr:nth-child(2) td:nth-child(4)")%>%
  html_text()
ownership

max.withdrawals.mgd <- URL_data%>%
  html_nodes("th~ td+ td")%>%
  html_text()
max.withdrawals.mgd

```


4. Convert your scraped data into a dataframe. This dataframe should have a column for each of the 4 variables scraped and a row for the month corresponding to the withdrawal data. Also add a Date column that includes your month and year in data format. (Feel free to add a Year column too, if you wish.)

>TIP: Use `rep()` to repeat a value when creating a dataframe.

>NOTE: It's likely you won't be able to scrape the monthly widthrawal data in chronological order. You can overcome this by creating a month column manually assigning values in the order the data are scraped: "Jan", "May", "Sept", "Feb", etc... Or, you could scrape month values from the web page...

5. Create a line plot of the maximum daily withdrawals across the months for 2022

```{r create.a.dataframe.from.scraped.data}
#4 

Month= c("01", "05", "09", "02", "06", "10", "03", "07", "11", "04", "08", "12")


df_dirty_durham<- data.frame("Month"=Month,
                             "Year"= rep(2022,12),
                             "max.withdrawals"= as.numeric(max.withdrawals.mgd))

df_dirty_durham<- df_dirty_durham%>%
  mutate(water.system=!!water.system.name,
         PWSID=!!pswid,
         Ownership=!!ownership,
         Date= my(paste0(Month,"-",Year)))
       
#5
ggplot(df_dirty_durham)+
  geom_line(aes(x=Date, y=max.withdrawals))+
  geom_point(aes(x=Date, y=max.withdrawals))+
  labs(y="max withdrawals (mgd)")+
  theme_bw()


```

6. Note that the PWSID and the year appear in the web address for the page we scraped. Construct a function using your code above that can scrape data for any PWSID and year for which the NC DEQ has data. **Be sure to modify the code to reflect the year and site (pwsid) scraped**.

```{r construct.a.scraping.function}
#6. 

scrape_it<- function(the_year, the_pwsid){
  scrapey_scrape<- read_html(
    paste0('https://www.ncwater.org/WUDC/app/LWSP/report.php?', 
           'pwsid=', the_pwsid, '&year=',the_year))
  
  
  water.system.tag<- 'div+ table tr:nth-child(1) td:nth-child(2)'
  pwsid.tag<- 'td tr:nth-child(1) td:nth-child(5)'
  ownership.tag<- 'div+ table tr:nth-child(2) td:nth-child(4)'
  max.withdrawal.tag<- 'th~ td+ td'
  
  Watersystem<- scrapey_scrape%>%
  html_nodes(water.system.tag)%>%
  html_text()
  
  Pwsid<- scrapey_scrape%>%
  html_nodes(pwsid.tag)%>%
  html_text()
  
  Ownership<- scrapey_scrape%>%
  html_nodes(ownership.tag)%>%
  html_text()
  
  MaxWithdrawal<- scrapey_scrape%>%
  html_nodes(max.withdrawal.tag)%>%
  html_text()
  
  df_<- data.frame("Month"=Month,
                             "Year"= rep(the_year,12),
                             "max.withdrawals"= as.numeric(MaxWithdrawal))
  
  df_<- df_%>%
  mutate(water.system=!!Watersystem,
         PWSID=!!Pwsid,
         Ownership=!!Ownership,
         Date= my(paste0(Month,"-",Year)))
  
  return(df_)
}

```

7. Use the function above to extract and plot max daily withdrawals for Durham (PWSID='03-32-010') for each month in 2015

```{r fetch.and.plot.Durham.2015.data}
#7 
durham2015<- scrape_it(2015,'03-32-010')
view(durham2015)

ggplot(durham2015)+
  geom_line(aes(x=Date, y=max.withdrawals))+
  geom_point(aes(x=Date, y=max.withdrawals))+
  labs(y="max withdrawals (mgd)")+
  theme_bw()


```

8. Use the function above to extract data for Asheville (PWSID = 01-11-010) in 2015. Combine this data with the Durham data collected above and create a plot that compares Asheville's to Durham's water withdrawals.

```{r fetch.and.plot.Asheville.2015.data}
#8 
Asheville2015<- scrape_it(2015, '01-11-010')

df_combined<-rbind(Asheville2015, durham2015)

ggplot(df_combined)+
  geom_point(aes(x=Date, y=max.withdrawals, color=water.system))+
  geom_line(aes(x=Date, y=max.withdrawals, color=water.system))+
  labs(y="Max Withdrawals (MGD)", 
       title = "Max Water Withdrawal Durham & Asheville 2015")+
  theme_bw()+
  theme(legend.position="left")


```


9. Use the code & function you created above to plot Asheville's max daily withdrawal by months for the years 2010 thru 2021.Add a smoothed line to the plot (method = 'loess').

> TIP: See Section 3.2 in the "10_Data_Scraping.Rmd" where we apply "map2()" to iteratively run a function over two inputs. Pipe the output of the map2() function to `bindrows()` to combine the dataframes into a single one. 

```{r fetch.and.plot.Asheville.multiyear.data }
#9 
the_year = seq(2010,2021)
the_pwsid = '01-11-010'

Asheville2010to2021<- the_year%>%
  map(scrape_it, the_pwsid ='01-11-010')%>%
  bind_rows()

ggplot(Asheville2010to2021)+
  geom_point(aes(x=Date, y=max.withdrawals))+
  geom_smooth(aes(x=Date, y=max.withdrawals))+
  labs(y="Max Withdrawals (MGD)", 
       title = "Max Withdrawal Asheville 2010-2021")+
  theme_bw()+
  theme(legend.position="left")


```

Question: Just by looking at the plot (i.e. not running statistics), does Asheville have a trend in water usage over time?
> Answer:
>From the plot, I found that Asheville has an increasing trend in water usage over time from 2015 and has the highest withdrawals in 10 years in 2021.
