---
title: "Assignment 2: Coding Basics"
author: "Miaojun Pang"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on coding basics.

## Directions

1.  Rename this file `<FirstLast>_A02_CodingBasics.Rmd` (replacing `<FirstLast>` with your first and last name).
2.  Change "Student Name" on line 3 (above) with your name.
3.  Work through the steps, **creating code and output** that fulfill each instruction.
4.  Be sure to **answer the questions** in this assignment document.
5.  When you have completed the assignment, **Knit** the text and code into a single PDF file.
6.  After Knitting, submit the completed exercise (PDF file) to Sakai.

## Basics, Part 1

1.  Generate a sequence of numbers from one to 30, increasing by threes. Assign this sequence a name.

2.  Compute the mean and median of this sequence.

3.  Ask R to determine whether the mean is greater than the median.

4.  Insert comments in your code to describe what you are doing.

```{r basics 1}
#1. 
seq(1,30,3)
three_sequence<-seq(1,30,3)
#created sequence beginning from 1, ending in 30, increasing by threes and assigned this sequence a name three_sequence


#2. 
mean(three_sequence)
#calculated mean

median(three_sequence)
#calculated median

#3. 
mean(three_sequence)>median(three_sequence)
##created a conditional statement asking R if mean is greater than median

```

## Basics, Part 2

5.  Create a series of vectors, each with four components, consisting of (a) names of students, (b) test scores out of a total 100 points, and (c) whether or not they have passed the test (TRUE or FALSE) with a passing grade of 50.

6.  Label each vector with a comment on what type of vector it is.

7.  Combine each of the vectors into a data frame. Assign the data frame an informative name.

8.  Label the columns of your data frame with informative titles.

```{r basics 2}
test_score<- c(99,66,33,69)
#test scores uses different numbers as the mode

student_name<-c("R", "Duke", "Nicholas", "EDA")
#student_name uses characters as the mode

test_pass_fail<-c(TRUE,TRUE,FALSE,TRUE)
#test_pass_fail uses characters as the mode

class_pass<-data.frame("student name"=student_name, "test score"=test_score, "test results"=test_pass_fail)

class_pass

```

9.  QUESTION: How is this data frame different from a matrix?

> Answer:
This data frame is different from a matrix. Because of it includes different modes while matrices does not have different modes. 

10. Create a function with an if/else statement. Your function should take a **vector** of test scores and print (not return) whether a given test score is a passing grade of 50 or above (TRUE or FALSE). You will need to choose either the `if` and `else` statements or the `ifelse` statement. 

11. Apply your function to the vector with test scores that you created in number 5.

```{r basics 3}

check_pass_fail <- function(scores) {
  for(score in scores) {
    if(score >= 50) {
      print(TRUE)
    } else {
      print(FALSE)
    }
  }
}

check_pass_fail(test_score)

```

12. QUESTION: Which option of `if` and `else` vs. `ifelse` worked? Why?

> Answer: "if" and "else" works for me. This function will iterate through each score in the test_score vector, printing TRUE if the score is 50 or above, and FALSE otherwise. This will provide a clear indication of pass/fail status for each student's score.